---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=TRUE}
library(dplyr)
myDf <- read.csv("activity.csv")
myDf <- na.omit(tbl_df(myDf))
head(myDf)
```


## What is mean total number of steps taken per day?

### 1. Make a histogram of the total number of steps taken each day

* prepares the data.
```{r echo=TRUE}
library(ggplot2)
stepsByDay <- 
    myDf %>%
    select(steps,date) %>%
    group_by(date) %>%
    summarise(steps = sum(steps))              

head(stepsByDay)
```

```{r echo=TRUE}
qplot(steps, data = stepsByDay, fill = date)
```


### 2. Calculate and report the mean and median total number of steps taken per day.

```{r echo=TRUE}
mean(stepsByDay$steps)
median(stepsByDay$steps)
```

* The mean number of steps taken per day is `r mean(stepsByDay$steps)`

* The median number of steps taken per day is `r median(stepsByDay$steps)`


## What is the average daily activity pattern?

### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
### and the average number of steps taken, averaged across all days (y-axis)

* Prepares the data
```{r echo=TRUE}
stepsByInterval <- aggregate(steps~interval, data=myDf,mean)
head(stepsByInterval)
```


* Makes the plot.
```{r echo=TRUE}
ga <- ggplot(data = stepsByInterval, aes(x=interval, y = steps))
ga <- ga + geom_line()
ga <- ga + xlab("5 minute interval")
ga <- ga + ylab("Average number of steps taken") 
ga <- ga + ggtitle("Average daily activity pattern")
ga

```

### Which 5-minute interval, on average across all the days in the dataset,
### contains the maximum number of steps?

```{r echo=TRUE}
stepsByInterval[which.max(stepsByInterval$steps),]
```

* The interval that contains the maximun number of steps is `r stepsByInterval[which.max(stepsByInterval$steps),]$interval`

## Imputing missing values

### 1. Calculate and report the total number of missing values in the dataset
### (i.e. the total number of rows with NAs)

* Read with missing values.
```{r echo=TRUE}
myDfWithNa <- read.csv("activity.csv")
head(myDfWithNa)
sum(is.na(myDfWithNa))
```

* The total number of missing values is `r sum(is.na(myDfWithNa))`

### 2. Devise a strategy for filling in all of the missing values in the dataset. The
### strategy does not need to be sophisticated. For example, you could use
### the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r echo=TRUE}
fillNaByInterval <- function (interval) {
    stepsByInterval[stepsByInterval$interval == interval,]$steps
}

```


### 3. Create a new dataset that is equal to the original dataset but with the
### missing data filled in.


```{r echo=TRUE}
myDfFilled <- myDfWithNa

for(i in 1:nrow(myDfFilled)) {
    if(is.na(myDfFilled[i,]$steps)) {
        myDfFilled[i,]$steps <- fillNaByInterval(myDfFilled[i,]$interval)
    }
}

head(myDfFilled)

```


### 4. Make a histogram of the total number of steps taken each day and Calculate
### and report the mean and median total number of steps taken per day. 

```{r echo=TRUE}

library(ggplot2)

stepsByDayFilled <- 
    myDfFilled %>%
    select(steps,date) %>%
    group_by(date) %>%
    summarise(steps = sum(steps))              

head(stepsByDayFilled)
```

### 1. Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
qplot(steps, data = stepsByDayFilled, fill = date)
```

* The mean number of steps taken per day is
```{r echo=TRUE}
mean(stepsByDayFilled$steps)
```


* The median number of steps taken per day
```{r echo=TRUE}
median(stepsByDayFilled$steps)
```



#### Do these values differ from the estimates from the first part of the assignment?
#### What is the impact of imputing missing data on the estimates of the total
#### daily number of steps?

* Is not so much different, the outpu is almost the same but with a litter diferences.


## Are there differences in activity patterns between weekdays and weekends?


### 1. Create a new factor variable in the dataset with two levels - "weekday"
### and "weekend" indicating whether a given date is a weekday or weekend
### day.

[Find the day of a week in R](http://stackoverflow.com/questions/9216138/find-the-day-of-a-week-in-r)
* For Sunday is 0 and saturday is 6.
```{r echo=TRUE}
myDfFilled$day <- 
    ifelse(
        as.POSIXlt(as.Date(myDfFilled$date)
        )$wday%%6==0,"weekend","weekday")


myDfFilled$day=factor(myDfFilled$day, levels=c("weekday","weekend"))

head(myDfFilled)
```



### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the
### 5-minute interval (x-axis) and the average number of steps taken, averaged
### across all weekday days or weekend days (y-axis). The plot should look
### something like the following, which was creating using simulated data:

* Prepares the data.
```{r echo=TRUE}
averagedAllWeek <- aggregate(steps~interval+day, data=myDfFilled, sum)
```

[reference facets ggplot2](http://docs.ggplot2.org/0.9.3.1/facet_grid.html)
```{r echo=TRUE}
ga <- ggplot(data = averagedAllWeek, aes(x=interval, y = steps))
ga <- ga + geom_line()
ga <- ga + facet_grid(day~.)
ga <- ga + xlab("5 minute interval")
ga <- ga + ylab("Number of steps") 
ga <- ga + ggtitle("Average daily activity pattern")
ga
```

